---
title: 'STA304 Final Report: Causal Inference Analysis on the Factor Immunization coverage that May Affect Life Expectancy'
author: "Qihui Huang, 1004932905"
date: "12/7/2020"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
bibliography: reference.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(skimr)
library(arm)
library(broom)
library(huxtable)
library(dplyr)
```

# Abstruct

# Key Words

Propensity Score, Causal Inference, Life expectancy, Immunization, Observational study, Health.

# Introduction

Nowadays, the concept of causality has become more and more popular, not only in our daily life, but also in the domain of Statistics studies. Causal inference examines the potential cause and effect relations, that is how the outcome will change when the cause varies. It has been widely used for analysis in health, social, behavioral and economic sciences. With the techniques like propensity score matching being introduced, a quasi-experimental method, it helps us create artificial control groups to estimate the relation of the possible cause and effect conveniently and minimize effect of confounding factors. Moreover, observational data extracted from WHO is used in this report instead of complete experiment data. As observational data is much more detailed and also a reliable source containing related rich information, more insights and creative observations/analysis can be performed here.

The main goal of this report is to find a causal relation between life expectancy and ea immunization coverage in countries around the world in a 15 years period. Life expectancy has always been an intriguing topic, and there have been lots of studies conducted in the past on factors affecting life expectancy, like personal income, morality, alcohol consumption, GDP and health care insurance etc. During COVID-19 pandemic, topics like immunization and vaccination have become prominent. As COVID vaccines are still in approval process, people all around the world are expecting the release of the vaccines that would lower down the number cases and prolong life. Therefore it is non-negligible when we consider the factors that may impose effects on life expectancy. This motivates us to explore the casual inference between the life expectancy and vaccination converge, and build a regression model based off multiple factors, considering data in a decade period all around the world. [@citeDataset]

The data set from [@citeKaggleWebsite] will be used in this analysis. We aim to find causal relations between factors with life expectancy. Other factors like countries economic condition, mortality and average education level will also be taken into consideration. In Methodology section(section 2), data cleaning is performed and model is built here. Propensity score matching is used here to discern whether individuals get vaccinated already and whether individuals die before the average age of population. Results of propensity score matching will be presented in Result section(section 3). Interpretation of the results, graphs, conclusion and further discussions will be included in the Discussion part(section 4).


# Methodology

## Data Description and Data Cleaning

The original data set is obtained from the World Health Organization (WHO) data repository. It contains detailed health related factors from 193 countries and corresponding economic status of each country from the United Nation website. It records people’s health status more than a decade, from 2000 to 2015. The data set has now been updated and reduced to 22 variables with 2938 rows by the user on Kaggle website, which contains mainly 5 categories: mortality factors, immunization related factors, economical factors, social/educational factors and personal habit factors.

I have modified data set sourced from the Kaggle website by performing some basic data cleaning steps, including mutating and choosing specific variables. There are in total 22 variables including both categorical and numerical, and these are all potential factors that may impose different levels of effect on life expectancy; 13 of them have been chosen and put into a new cleaned data set, like *life expectancy*, *country*, *BMI* and different types of vaccines etc.. These variables are more representative and clean compared to other factors, and will be easily manipulated in the later analysis. For convenience's sake, especially in matching propensity scores, a new binary variable named *vaccine_coverage_over_ave* been created, measuring whether a country’s immunization coverage exceeds the world’s average or not. This is simply done by adding all general vaccines coverage rate together for each country, then compared each of them with the mean of the world vaccine coverage. In this case, 1 represents the country has met the world’s average immunization coverage, and 0 represents the country has not met the criteria here yet. Moreover, renaming variables and removing rows with NA or invalid numbers are done in this part as well.


#### A glimpse of cleaned data set:

```{r,echo=FALSE,message=FALSE}
#rename and select predictors/variables for later usage.
raw_data <- read_csv("life_expectancy.csv")
cleaned_data <- raw_data %>%
  rename(country = `Country`) %>%
  rename(status = `Status`) %>%
  rename(population = `Population`) %>%
  rename(life_expectancy = `Life expectancy`) %>%
  rename(total_expenditure = `Total expenditure`) %>%
  rename(Hepatitis_B = `Hepatitis B`) %>%
  rename(HIV_AIDS = `HIV/AIDS`) %>%
  rename(healthcoverage_percent = `percentage expenditure`) %>%
  rename(adult_mortality = `Adult Mortality`) %>%
  filter(Schooling != 0.0) %>%
  dplyr :: select(country, status, life_expectancy, adult_mortality, Alcohol, healthcoverage_percent, Hepatitis_B, BMI, Polio, total_expenditure, Diphtheria, GDP, Schooling)

cleaned_data <- na.omit(cleaned_data)

# Mutate a binary treatment immunization coverage
cleaned_data <- cleaned_data %>%
  mutate(vaccine_total = Polio+Diphtheria+Hepatitis_B) %>%
  mutate(vaccine_coverage_over_ave = ifelse(vaccine_total >= mean(vaccine_total), 1, 0))

knitr::kable(cleaned_data[1:15, 2:7])
```

## Model

The propensity score (probability of being treated) will be estimated by constructing a logistic regression model at first. Our treatment is immunization coverage in a country, and life expectancy is our outcome of interest, so that propensity score matching will be for the immunization coverage propensity. Nearest neighbor matching will be used next to find the closest match in the untreated group, keeping only the rows that are matched. Lastly, a matched sample will be run for common multiple regression analysis.

Firstly, logistic regression model is carried out, where the treatment immunization coverage is now presented as outcome based on different co-variates in our dataset. In this step, we think that predictors like countries' status, alcohol consumption, adult morality, GDP etc. are able to explain whether a country is treated or not. The model gives out the probability of the treatment. Propensity score is obtained here and added to the data set, which is exactly the probability of a country's vaccine coverage over world's average.

Summary coefficient table of fitted logistic regression model between treatment vaccine coverage and other factors

Intercept/variables | Estimates | P-value 
-----------|:-----------: |:-----------: |
intercept  | -2.0880 | 4.81e-06 |   
statusDeveloping  | -0.6484 | 0.0117 |
adult_mortality  | -0.0013 | 0.0049 |
Alcohol  | -0.0189 | 0.3301 |
total_expenditure  | 0.0756 | 0.0023 |
GDP  | 9.671e-05 | 9.60e-07 |
BMI  | 0.0016 | 0.6362 |
Schooling  | 0.7556 | < 2e-16 |
healthcoverage_percent  | 0.0007 | 4.84e-09 |

```{r,echo=FALSE, message=FALSE}
#logistic model
propensity_score <- glm(vaccine_coverage_over_ave ~ status+adult_mortality+Alcohol+total_expenditure+GDP+BMI+Schooling+healthcoverage_percent,
                        family = binomial,
                        data = cleaned_data)
```

```{r,echo=FALSE, message=FALSE}
# add forecast to our data set
cleaned_data <- augment(propensity_score,
                        data = cleaned_data,
                        type.predict = "response") %>%
  dplyr::select(-.resid, -.std.resid, -.hat, -.sigma, -.cooksd)

cleaned_data <- cleaned_data %>%
  arrange(.fitted, vaccine_coverage_over_ave)

```


```{r,echo=FALSE, message=FALSE}
cleaned_data$treated <- ifelse(cleaned_data$vaccine_coverage_over_ave == 1, 1, 0)
cleaned_data$treated <- as.integer(cleaned_data$treated)
matches <- arm::matching(z = cleaned_data$treated,
                         score = cleaned_data$.fitted)
cleaned_data <- cbind(cleaned_data, matches)


cleaned_data_matched <- cleaned_data %>%
  filter(match.ind != 0) %>%
  dplyr::select(-match.ind, -pairs, -treated)

head(cleaned_data_matched)

propensity_score_regression <- lm(life_expectancy~status+adult_mortality+Alcohol+total_expenditure+GDP+BMI+Schooling+healthcoverage_percent+vaccine_coverage_over_ave, data = cleaned_data_matched)

summary(propensity_score_regression)
knitr::kable(huxtable::huxreg(propensity_score_regression))
```




# Github Repo

https://github.com/heyhaileyhhh/304finalreport_qh.git

# Appendix

Summary table of logistic regression model output between the treatment vaccine coverage and other co-variates
```{r,echo=FALSE,message=FALSE}
summary(propensity_score)
```

Summary table of logistic regression model output between the outcome life expectancy with propensity score added
```{r,echo=FALSE,message=FALSE}
summary(propensity_score_regression)
```


# References


